### 2016-07-26 src/ack.tex
Renamed the ICE paper.
Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

\section{Acknowledgements}
The authors are grateful for the assistance and support of the following people
and institutions without which this work would not have been possible.

\textbf{FIXME!}

This work has been supported by the US Department of Energy, Offices of Nuclear
Energy (DOE-NE) and Energy Efficiency and Renewable Energy (DOE-EERE), and by
the ORNL Undergraduate Research Participation Program, which is sponsored by
ORNL and administered jointly by ORNL and the Oak Ridge Institute for Science
and Education (ORISE). This work was also supported in part by the Oak Ridge
National Laboratory Director's Research and Development Fund. ORNL is managed by
UT-Battelle, LLC, for the US Department of Energy under contract no.
DE-AC05-00OR22725. ORISE is managed by Oak Ridge Associated Universities for
the US Department of Energy under contract no. DE-AC05-00OR22750.

\textbf{LDRD ACK?!}