### 2016-07-26 src/main.tex

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass[]{elsart}

\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}

\begin{document}

\begin{frontmatter}

\title{The Eclipse Integrated Computational Environment}

\author{Jay Jay Billings$^{1,2}$}
\ead{billingsjj@ornl.gov}
\author{Alexander J. McCaskey$^{1}$}
\author{ADD ALL OTHER AUTHORS}

%%%%% ----- FIXME! Add all ICE & NiCE contributors

\address{$^1$Computer Science and Mathematics Division, Oak Ridge National
Laboratory, Oak Ridge, TN 37830, USA}
\address{$^2$The Bredesen Center for Interdisciplinary Research and Graduate
Education, University of Tennessee, Knoxville, TN 37996-1200, USA}

%----------------\input{abstract}

\begin{keyword}
\sep 
\sep 
supercomputing

\vspace{1ex}

\PACS
07.05.Tp % Computer modeling and simulation

\end{keyword}

\end{frontmatter}

\input{intro}
%--------\input{background}
%--------\input{design}
%--------\input{comparisons}
%--------\input{future}
%--------\input{availability}
\input{conclusions}
\input{ack}

\bibliographystyle{unsrt}
\bibliography{bib}

\end{document}