### 2017-01-01 src/content.tex
Committing first draft of ICE paper.
Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

\section{Motivation and Significance}\label{motivation-and-significance}

High performance modeling and simulation software is hard to use. Much
of the challenge comes from the inherent nature of the science under
consideration, but perhaps just as much of the problem comes from
focusing very heavily on raw compute performance. Focusing on
performance is understandable, given the field, but it leaves the rest
of us to figure out how to actually use the software in a way that
scales for a much larger set of users, many of whom may be novices or
migrate between software packages regularly.

** REMOVE or merge with text below!**

In previous work, Billings et. al., discovered through requirements
gathering interviews that many of the difficulties using high
performance modeling and simulation software fall broadly into five
distinct categories, \cite{billings_cbhpc}. This includes input
generation and preprocessing (also called ``model setup''); job
execution and monitoring; postprocessing, visualization and data
analysis; data management; and customizing the software. There are many
tools that address these problems individually, but the same work found
that the excess number and specialization of these tools also contribute
to the learning curve.

Efforts to address these five issues typically fall in with general
purpose scientific workflow tools like Kepler, \cite{kepler}, or are
reduced to myopic tools that satisfy some set of requirements for a
single piece of software or platform. That is, the proposed solution to
this problem is often to shoe-horn it into existing workflow tools that
are so general that they focus on nothing in particular or to ignore the
general problem entirely and deploy a completely tailored solution for
the given application. These are opposing extremes, but a
middle-of-the-road solution is also possible. A workflow engine could be
developed that limits its scope to High-Performance Computing (HPC) and
to the set of possible workflows that come from the previously mentioned
five activities. A rich enough Application Programming Interface (API)
could be exposed so that highly customized solutions could still be made
based on this limited workflow engine with only a relatively minor
amount of additional development required.

It is not clear that one of these solutions is better than the others.
Practical requirements will ultimately dictate which way projects go.
This work considers a middle ground solution and presents the Eclipse
Integrated Computational Environment (ICE) as proof that it is possible
to create such a system. Specifically, we show * an architecture for
such a workflow system that satisfies the model of workflows presented
below in an extensible way. * that such a system can be cross-platform
for workstations and simultaneously deployable on the web. * that smart
design decisions enable not only authoring code for simulation software,
but also make it possible for the system to extend itself, thereby
enabling heavy customization.

\textbf{FIXME! Needs to transition smoothly!}

\subsubsection{Workflow Model}\label{workflow-model}

Computational scientists perform a variety of tasks with respect to
modeling and simulation that can be abstracted into a lightweight
theoretical framework broadly based on five high-level
\emph{activities}: creating input, executing jobs, analyzing results,
managing data, and modifying code. The same computational scientist
would most likely find these activities difficult for all codes with
which they lack experience, whereas with their own codes or those with
which they are most familiar these tasks may be so easy that they are
taken for granted. Any particular combination of these activities across
one or more scientific software packages or codes results in a unique
\emph{workflow}. Such a workflow is normally, but not always, requested
by a human user and orchestrated by a \emph{Workflow Management System
(WMS)}.

The most obvious workflow for any individual simulation code or
collection of codes is to string the activities together: the user's
workflow is to create the input, launch the job, perform some analysis
and manage their data, possibly modifying their code in the process.
There are many other combinations though, such as re-running jobs with
conditions or modifications or analyzing someone else's data. (See
footnote 1.)

\textbf{Creating input} is the process of describing the physical model
or state of a system that will be simulated. This could include creating
an input file or file(s), but could also include external calls to a
process to configure a running program. In most situations a
computational scientist will modify existing input or create new input
from a template. ``Input'' generally includes runtime parameters for the
simulation framework (tolerances, etc.); configuration options (data
locations, output locations, module configurations, etc.); material
properties for the materials that will be simulated; and a
discretization of the simulation space (mesh, grid, particle
distribution, etc.). The collection of all required input can be quite
large and may go by many names such as ``input set,'' ``input package,''
``problem,'' or, simply, ``input.'' Often the set of input files will be
described in a ``main'' input file that acts as a kind of manifest,
describing and providing links to all necessary information for the
given problem.

In this work, it should be assumed unless otherwise noted that ``input''
refers to the entire set of input, not a single file.

\textbf{Executing jobs} or ``running the workflow'' in this context is
the process of performing some calculation based on the input with a
simulation code or framework. Modeling and simulation codes are normally
run locally for small jobs or development. Large simulations typically
require equally large hardware resources that are utilized remotely
through Secure Shell (SSH) connections or similar protocols. Remote
execution requires moving the input in advance of the execution and
copying or moving the output to the user's machine, if possible and
desirable. (In many cases the output cannot be moved because it is too
large.)

Local and remote jobs are often monitored in one or more ways to
ascertain the status of job. This can include simply checking whether or
not execution has finished or monitoring the output of individual
quantities to examine the calculation state. The latter is often used to
detect calculation errors that will result in incorrect results. If such
results are found, the job is typically cancelled (``killed'') to save
compute resources and re-run later.

In this work, it should be assumed unless otherwise noted that
``executing a job'' includes monitoring that job in one or more ways,
possibly including real-time updates to visualizations.

\textbf{Analyzing results} includes executing special jobs to transform
data in one or more prescribed ways and producing artifacts with
scientific significance from the transformed data. This may include, for
example, post-processing results and visualizing the new data with
dedicated visualization tools. For many types of scientific computing
this includes viewing the results of a simulation on a mesh or grid and
extracting publication quality images or movies from that data. Other
cases may include analyzing results in preparation for follow
simulations or performing feature extraction, classification or
activities for machine learning and data mining.

While this has many similarities to executing a job, it is distinctly
different because the activity changes focus to satisfy the needs of a
human operator. Simple data reduction where the exact reduction is known
certainly qualifies as executing a job, but analysis of modeling and
simulation results is far from simple data reduction. It is generally a
far more interactive process for computational scientists.

\textbf{Managing data} includes moving, copying, storing, sharing or
otherwise interacting with data for or from simulations. This activity
is the most pervasive of all the activities because each of the others
requires interacting with data in some way. In many cases though, data
is still managed its own purposes, without performing a simulation,
generating new input, or analysing results. Examples include archiving
data, packaging data for publications, and updating values manually
(often in light of new information from publications).

\textbf{Modifying code} is not typically considered a part of a
scientific computing workflow. However, modeling and simulation use
cases often require users to explicitly modify code before execution,
such as with the computational fluid dynamics code Nek5000,
{[}Nek5000{]}; or to issue special re-build instructions. Many
computational scientists consider ``their workflow'' to be re-running
software after modifications for purely exploratory purposes. This may
be required if the model the author is manipulating can not be
configured as part of the input directly but can be easily manipulated
by hand.

\subsubsection{Comparison to Other
Models}\label{comparison-to-other-models}

This model of workflows differs significantly from those of similar
efforts in workflow science because it defines workflows in terms of
activities. Others in the literature define a workflow as a collection
of computing processes. For example, Yu and Buyya define grid workflows
as ``a collection of tasks that are processed on distributed resources
in a wel-defined order to accomplish a specific goal,'' {[}Yu, 2005{]}.
Others such as Pizzi et al. subscribe to similar definitions, {[}Pizzi,
2015{]}. This ``process'' view is acceptable where the workflow is
static and does not require additional human input or ``human in the
loop'' behavior after the all the initial human input is provided. That
is, a ``process-oriented'' definition is acceptable where all human
input is provided in advance. However, workflows within ICE are fully
interactive with regular call-backs to humans. It is simpler to discuss
``activities'' than it is to create a distinction between ``human
processes'' and ``computer processes.'' Focusing on ``activities'' over
processes (human or computer) also has the benefit of removing concrete
elements such as hardware properties or software details that distract
from details of workflows and WMSes per se. That is, considerations such
as memory usage and raw performance are important, but questions about
the abstract workflow or what the WMS should do are far more important.

\section{Software Description}\label{software-description}

ICE was specifically created to address the workflow model described
above, which is to say that it was created specifically to address
hands-on workflows for computational scientists. Users download and
execute ICE locally and it orchestrates workflows locally or remotely as
required. It provides a comprehensive workbench for modeling and
simulation that includes tools for workflows, visualization, data
management, and software development.

\subsection{Software Architecture}\label{software-architecture}

Workflows and tasks in ICE are not explicitly treated as trees -
Directed Acyclic Graphs, (DAGS) - as is common with grid workflow tools,
{[}Yu, 2005{]}. Instead, ICE's design is inspired by Representational
State Transfer (REST), {[}Fielding, 2000{]}.

Users initiate requests to create, edit, update or delete workflows from
the \emph{ICE Client} (the workbench). The list of available workflow
that can be created is provided dynamically to the \emph{ICE Client} by
the \emph{ICE Core}, which acts as a kind of server. This information is
provided dynamically since it often changes as run time based on the
configuration of available workflow components in the registry and on
persisted workflows that users have saved in their workspace.
Information about workflows are provided from the Core to the Client
through \emph{stateless Forms} that describe the workflow and provide
all the necessary information to understand \emph{what} should process
the workflow, (but not \emph{how} it should be processed). Once users
modify the description of the workflow in the Form to provide their
specific details, the Client dispatches a request to the Core to modify
or process the workflow. The Core then uses the information in the Form
to perform a service lookup for \emph{what} should process the workflow.

Workflows in ICE are encoded in and processed by \emph{Items} and each
workflow type is a subclass of Item. Items are registered dynamically
through a service registry in ICE, (see section \ref{framework}), and
provide the Forms for the ICE Core and Client. Since ICE's design is
highly object oriented, it is easiest to think of the Item class as a
description of an abstract workflow and an Item object (an instance of
the class) as a concrete workflow with all required execution details
provided by its Form.

Individual components of workflows, i.e. - workflow ``tasks'' or
``nodes'' - are either encoded directly in the workflow's subclass of
Item or provided as \emph{Actions} that are dynamically registered with
an \emph{Action Factory} and obtained at runtime. Table 1. describes the
differences between Items, Actions, and Forms. (See footnote 2.)

Table 1: Class Descriptions for Items, Forms and Actions

\begin{longtable}[c]{@{}lll@{}}
\toprule
Class & Class Description & Object Description\tabularnewline
\midrule
\endhead
Item & Java class with code to execute an abstract workflow. Provides a
Form. & Concrete workflow executor.\tabularnewline
Form & Description and template of the data needed for the Item to
process the workflow. & User-modified workflow data.\tabularnewline
Action & Java class for executing a specific task in the workflow. Used
by Items. & Concrete workflow task executor.\tabularnewline
\bottomrule
\end{longtable}

In addition to running as a desktop workbench, ICE can be run as a
headless web server with a remote service interface and a web
Application Programming Interface (API). The web API is also used as the
primary means of providing real-time feedback and monitoring support in
ICE.

\subsubsection{Item States}\label{item-states}

All Items in ICE are finite state machines where the states represent
the abstract state of the workflow. For example, when an Item is first
created, it enters the ``Form Ready'' state to indicate that it could,
in theory, be processed after a user reviews it. After that review it
enters the ``Ready to Process'' state before and the ``Processed'' state
after it is processed. There are several additional states for errors.

This design is very important. First, it means that all workflows in
ICE, regardless of their actual details and function, can only behave in
a specific set of known, predictable ways. This greatly simplifies the
way that the Core manages and interacts with Items. Second, it
explicitly delineates which workflows can be executed from those that
must still receive some configuration because it formalizes state and
error checks. Finally, it allows developers implementing Items and
Actions to specify by contract what is required before proceeding to the
next task, processing the workflow, or declaring a successful execution.

\subsubsection{Persistence and
Workspaces}\label{persistence-and-workspaces}

ICE persists permanent copies of Forms to disk when workflows are
created and modified in a special directory called a \emph{workspace}.
Workspaces can contain projects, folders and files, including data,
code, input and output. ICE automatically manages local and remote (or
even local \emph{to} remote) transfers of data files when executing
workflows if the files are detected in the same directory of the
workspace as the workflow itself. For example, when executing a remote
job, ICE will automatically move the input file if it is specified in
the workflow and available in the project directory of the workspace.
Likewise, if output is small enough, ICE will automatically move it back
to local directory. By conventional, all paths in ICE's Forms are
relative to the workspace root path. Workspace directories are specified
by the user.

ICE handles persistence using a \emph{persistence provider}. The only
default persistence provider uses JAX-RS, {[}Burke, 2009{]}, to write
Forms to XML. In principle other persistence providers could replace the
XML-based provider since it is registered as a dynamic service, and in
the past a JAX-P based provider has been used.

\subsubsection{Extending the Framework to Add Custom
Workflows}\label{extending-the-framework-to-add-custom-workflows}

ICE is an Eclipse Rich Client Platform (RCP) application, {[}McAffer 2,
2010{]}. It has a plugin architecture based on Equinox, the reference
implementation of the Open Service Gateway Initiative (OSGi) framework
specification, {[}McAffer, 2010{]}. It uses over 1200 additional
packages from the Eclipse ecosystem to provide services such as language
support and visualization to users. Each unique element of ICE described
in this work, including the client, core, all Items, data structures and
many others are provided as plugins to Equinox. Most plugins are
dynamically managed and provided as services that can be obtained from a
service registry. All file I/O in ICE, with only a few exceptions,
interacts with the RCP's Resources plugin. This includes remote
resources that are managed with the Eclipse Parallel Tools Platform,
{[}Tibbitts, 2009{]}.

It is not possible to graphically create new Items in ICE. Users can
only use the workflow Items provided with their version of ICE unless
they provide their own plugin to the framework. This can seem like an
onerous task to the novice user especially if they do not know Java, but
this design was chosen because it is, in the opinion of the authors, a
far more sophisticated and sustainable way to create workflows than
through wiring diagram.

ICE provides built-in development tools to auto-generate ``stubs'' of
plugins that can then be installed into the framework. This
``self-hosting'' makes it possible for new users to create complex,
sophisticated workflows very quickly because they are not required to
know how to interact with the framework. Furthermore, ICE provides a
rich API, documentation, tutorials, and tools to further simply the
process.

There are many situations where graphically configuring workflows is
unacceptable, such as when a very large number of workflows will be
executed or the type of information required is very fine-grained. In
these cases, it is often necessary to provide scripts to the workflow
management system. ICE includes the Eclipse Advanced Scripting
Environment, {[}Pontesegger, 2017{]}, (EASE), for scripting because it
provides a native way to script Eclipse RCP projects in Javascript,
Jython, and Python. This also makes it possible to extend the
environment by adding Items in these languages.

\subsection{Software Functionalities}\label{software-functionalities}

The most important function of ICE is to serve as an easily extended
workflow management system for computational scientists in support of
the activities described above. In practice, it is most often used as a
combination of workflow management system and development environment
since it contains a significant amount of Eclipse's software development
tooling in addition to the workflow tools. The project has many users,
but it is most heavily used by the development team to support workflow
science and software development for energy science projects. The
development team regularly uses the platform to quickly deploy new
domain-specific workbenches in a matter of hours for small collections
of workflows that are easy to encode.

Outside of the development team, ICE is commonly deployed as a
sophisticated user environment for computational science projects (see
``Impact'' below) and as a visualization tool. ICE originally a
significant amount of visualization support, but at the request of users
in the community that support was ``spun off'' in early 2016 as a
separate visualization project, the Eclipse Advanced Visualization
Project, EAVP, {[}EAVP, 2009{]}.

\section{Illustrative Examples}\label{illustrative-examples}

The role that ICE plays as a workflow tool is best illustrated by the
various ways that it has been deployed.

\subsection{Virtual Battery
Simulations}\label{virtual-battery-simulations}

Pannala et al. developed a Virtual Integrated Battery Environment (VIBE)
as part of their research into safety and performance characteristics of
lithium ion batteries, {[}Pannala, 2015{]}. VIBE provides ICE to users
as part of its distribution. New workflows were added to ICE to create
multiple types of input, configure the simulation software, and launch
simulations of virtual batteries. Interactive 3D visualizations of the
results were embedded in the launcher so that users would quickly find
their results. Figure 1 shows the results from a temperature
distribution of a prismatic cell battery simulated during discharge.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{images/vibe_20151016.png}
\caption{vibe}
\end{figure}

VIBE 1.0 is available as a virtual machine in which the simulation
software and ICE are provided side by side. More recent efforts for VIBE
1.1 by the VIBE development team include providing the simulation
software in Docker containers so that users can download the latest,
native version of ICE for their machine while simultaneously benefiting
from a smaller virtual machine for the simulator.

\subsection{Multiphysics Simulations with
MOOSE}\label{multiphysics-simulations-with-moose}

The MOOSE Framework is a powerful, easy to use multiphysics framework
developed at the Idaho National Laboratory, {[}Gaston, 2009{]}. ICE
provides workflow tools for MOOSE as well as specialized class
generation utilities for developing custom MOOSE kernels. Many of the
MOOSE tools in ICE were developed closely with the MOOSE team to closely
reproduce various aspects of MOOSE's user interface, Peacock. Figure 2
shows an example of the ICE workbench for a simple structural mechanics
problem solved with the MOOSE framework, {[}McCaskey, 2016{]}.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{images/ice-moose.png}
\caption{moose}
\end{figure}

There are over three hundred MOOSE-based applications and it is very
easy to create new MOOSE applications. The ICE development team uses ICE
and MOOSE to quickly solve energy science problems with high-performance
computing resources and deploy domain specific workbenches. ICE includes
support for automatically downloading and building MOOSE from its GitHub
project, which then makes it possible to immediately begin developing
complex multiphysics applications using the Eclipse C Development Tools,
which are pre-installed in ICE to facilitate this interaction. Once the
new MOOSE-based application is built, it will automatically work with
the MOOSE workflow tools in ICE, although developers can also create
customized workflow tools as needed.

\subsection{Binder Jet Modeling}\label{binder-jet-modeling}

Solid-state sintering of parts printed using binder jetting
significantly increases part strength by decreasing the part porosity
and eliminating voids. The downside of this process is that it causes
the part to shrink and warp. An ideal near-net-shape process that
combined binder jetting with solid-state sintering would thus consider
warpage and deformation due to sintering in the design phase of the
part. Figure 3 shows an ICE-based workbench for perform simulations of
this process with visualizations of the pre- and post-simulation
properties of a central body with eight cantilevers. The primary
deformation in this type of geometry is bending or drooping of the
cantilevers, and ICE currently calls a custom MOOSE-based application
(which was written in ICE as described above) for simulating the
deformation of the cantilevers.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{images/ice-bjm.png}
\caption{bjm}
\end{figure}

The full application for this work is expected to be released near the
end of 2017.

\subsection{Neutron Reflectivity}\label{neutron-reflectivity}

ICE includes a small utility for simulating neutron reflectivity and
comparing the results to data, {[}Billings, 2015{]}. This was developed
in collaboration with colleagues at the Spallation Neutron Source to
replace an older utility written in Visual Basic and distributed via
Excel macros. The newer utility in ICE is capable of processing a single
workflow and is shown in figure 4.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{images/reflectivity-screenshot.png}
\caption{reflectivity}
\end{figure}

\subsection{Quantum Computing}\label{quantum-computing}

As quantum computing grows, so to does the need for sophisticated
software that can utilize quantum hardware or perform calculations on
simulated quantum hardware. Humble et al. created a simulator for
adiabatic quantum computers where workflows were added to ICE to support
interaction with the simulator, but also to process large sets of
quadratic binary optimimzation problems, {[}Humble, 2014{]}. Figure 5
shows the workbench for this project.

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth]{images/jaded.png}
\caption{qc}
\end{figure}

ICE also supports other quantum computing projects, including the
eXtreme-scale ACCelerator (XACC) programming framework,
{[}ORNL-QCI/XACC, 2016{]}. XACC is a programming framework for
extreme-scale, post-exascale accelerator architectures that integrates
alongside existing conventional applications.

\subsection{Nuclear Energy}\label{nuclear-energy}

There are many examples of ICE's role in modeling and simulation
projects for nuclear energy, but readers are referred to the work in
{[}Billings, 2015{]} for an extreme example of the level of
customization that is possible in ICE. Support for the ``Reactor
Analyzer'' was dropped in ICE 2.1.8, but it demonstrated ICE's ability
to integrate many different nuclear energy tools for complex analysis.

\section{Impact}\label{impact}

The impact of software tools such as ICE are hard to quantify. However,
there are several places where ICE has significantly assisted its
development team and others.

One pressing area of interest and impact is that of interoperability
between workflow systems. There is significant prior art in combining
large workflow management systems, such as that Mandal et al.,
{[}Mandal, 2007{]}, but the end goal of gaining the greatest advantage
by using the best capabilities from multiple systems remains elusive.
ICE's unique perspective on workflows and its well-defined API make it
possible to integrate multiple systems in a straightforward way. This
allows it to connect to other workflow environments, such as Triquetrum,
quite easily, {[}Brooks 1, 2015{]}. Triquetrum, (like Kepler in
{[}Mandal, 2007{]}), is a Ptolemy-based workflow engine, {[}Brooks 2,
2015{]}. Several of the authors of this paper are using it to
investigate these issues.

It is widely known that tools that make researchers more productive tend
to improve the pursuit of existing research. The high extensibility of
ICE and the tools that it combines from the larger Eclipse ecosystem
have made it possible for researchers on the development team to quickly
deploy new simulation environments for problems, such as the binder jet
modeler and reflectivity tools mentioned earlier. Other tools create
with ICE may not radically invent something new, but they tend to
streamline interactions with those tools. Many ICE users (and certainly
the development team!) have experienced improvements in software
development activities because of the tools that ICE provides or learned
new technologies because access to new tooling was a simple as
installing more plugins through the Eclipse Marketplace.

The ICE development team does not track ICE's user base, as useful as
that would be, because of the extra work involved. However, various
sources such as the VIBE mailing list, ICE's own mailing lists, and
website download statistics suggest that ICE has been used by over 350
people at one time or another and currently has about twenty
``superusers,'' including the development team.

A new cloud based development tool based on ICE is under development by
RNET LLC out of Dayton, Ohio in response to an SBIR award from the U.S.
Department of Energy. This web based version of ICE will continue ICE's
support for nuclear energy and integrate with cloud services such as
Amazon Web Services and Oak Ridge National Laboratory's Compute and Data
Environment for Science (CADES). Additionally, while ICE has not led to
the development of ``spin-off'' companies, ICE source code has been used
in two spin-off projects: EAVP (mentioned earlier) for advanced
visualizations and the Eclipse January project for scientific data
structures, {[}Graham, 2016{]}. ICE was also one of the founding
projects of the Eclipse Science Working Group.

\section{Sample Code, Tutorials and Other
Resources}\label{sample-code-tutorials-and-other-resources}

The quintessential resource for information on ICE is the projects web
page, {[}ICE, 2016{]}. The ``Resources'' menu includes links to detailed
tutorials and documentation of multiple types. Examples of how to create
new workflow Items are available at
https://github.com/eclipse/ice/tree/master/org.eclipse.ice.demo.
Examples of how to use the scripting engine (in this case for neutron
reflectivity) are available at
https://github.com/eclipse/ice/tree/master/examples. ICE includes an
extensive test suite of unit, integration, and UI tests and these are
also excellent examples of how to work with the platform.

\section{Conclusions}\label{conclusions}

Workflows for computational scientists in modeling and simulation differ
greatly from experimentalists or those who primarily interact with
grid-based workflow management systems. This work presented the Eclipse
Integrated Computational Environment (ICE) and described how its has
been used to address interdisciplinary problems in modeling and
simulation for energy science. One interesting avenue for future
exploration is coupling or integrating ICE with other workflow tools
such as Aiida, Triquetrum, Kepler and Pegasus. It remains as abundantly
clear now as it was in Mandal's 2007 paper that the best of all workflow
management tools will be needed to address high-performance computing
problems at the exascale and beyond.

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

The authors are grateful for the assistance and support of the following
people and institutions without which this work would not have been
possible. This includes many people who directly contributed to the
project, either in its early days as ``NiCE'' or once it moved to
Eclipse, including: Ronald Allen, Andrew Belt, Tim Bohn, David E.
Bernholdt, Erica Grant, Mike Guidry, Forest Hull, Eric J. Lingerfelt,
Sebastien Jourdain, JiSoo Kim, Allison Koenecke, Fangzhou Lin, Greg
Lyon, Tony McCrary, John M. Hetrick III, Elizabeth Piersall, Neeti
Pokhriyal, Adrian Sanchez, Claire Saunders, Nick Stanish, Matthew Wang,
Scott Wittenberg. The support of both the NEAMS and CASL programs is
greatly appreciated. The authors would like to acknowledge the special
contribution of the Eclipse Foundation, the Eclipse Community, the
Eclipse Science Working Group and our many colleagues who use and
contribute to open source projects in the Eclipse ecosystem.

Finally, the development team is especially grateful to Barney Maccabe,
David Pointer and John Turner of Oak Ridge National Laboratory for their
endless support and advocacy for this work.

This work has been supported by the US Department of Energy, Offices of
Nuclear Energy (DOE-NE) and Energy Efficiency and Renewable Energy
(DOE-EERE), and by the ORNL Undergraduate Research Participation
Program, which is sponsored by ORNL and administered jointly by ORNL and
the Oak Ridge Institute for Science and Education (ORISE). This work was
also supported in part by the Oak Ridge National Laboratory Director's
Research and Development Fund. ORNL is managed by UT-Battelle, LLC, for
the US Department of Energy under contract no. DE-AC05-00OR22725. ORISE
is managed by Oak Ridge Associated Universities for the US Department of
Energy under contract no. DE-AC05-00OR22750.

\section*{Required Metadata}\label{required-metadata}
\addcontentsline{toc}{section}{Required Metadata}

\section*{Current code version}\label{current-code-version}
\addcontentsline{toc}{section}{Current code version}

--------- \textbar{}
:-----------------------------------------------------------------:
\textbar{} :-------------------------------------------------------- :
\textbar{}\\C1 \textbar{} Current Code Version \textbar{} `next'
\textbar{}\\C2 \textbar{} Permanent link to code/repository used for
this code version \textbar{} https://github.com/eclipse/ice/tree/next
\textbar{}\\C3 \textbar{} Legal Code License \textbar{} Eclipse Public
License 1.0 \textbar{}\\C4 \textbar{} Code versioning system use
\textbar{} Git \textbar{}\\C5 \textbar{} Software code languages, tools,
and services used \textbar{} Java, OSGi, Eclipse Rich Client Platform,
and Maven \textbar{}\\C6 \textbar{} Compilation requirements, operating
environments \& dependencies \textbar{} Java 1.8 or greater, Maven, and
an internet connection for dependencies \textbar{}\\C7 \textbar{} If
available Link to developer documentation/manual \textbar{}
https://wiki.eclipse.org/ICE \textbar{}\\C8 \textbar{} Support email for
questions \textbar{} ice-dev@eclipse.org \textbar{}

\section*{Current executable software
version}\label{current-executable-software-version}
\addcontentsline{toc}{section}{Current executable software version}

--------- \textbar{}
:-----------------------------------------------------------------:
\textbar{} :-------------------------------------------------------- :
\textbar{}\\S1 \textbar{} Current Code Version \textbar{} 2.2.1
\textbar{}\\S2 \textbar{} Permanent link to executables of this version
\textbar{}
http://www.eclipse.org/downloads/download.php?file=/ice/builds/2.2.1/
\textbar{}\\S3 \textbar{} Legal Software License \textbar{} Eclipse
Public License 1.0 \textbar{}\\S4 \textbar{} Computing
platforms/Operating Systems \textbar{} Windows (32/64-bit), Mac OS/X,
Linux (32/64-bit) \textbar{}\\S5 \textbar{} Installation requirements \&
dependencies \textbar{} Java 1.8 or greater \textbar{}\\S6 \textbar{} If
available, link to user manual - if formally published include a
reference to the publication in the reference list \textbar{}
https://wiki.eclipse.org/ICE \textbar{}\\S7 \textbar{} Support email for
questions \textbar{} ice-users@eclipse.org \textbar{}

\section{References}\label{references}

Yu, Jia, and Rajkumar Buyya. ``A Taxonomy of Workflow Management Systems
for Grid Computing.'' Journal of Grid Computing 3.3--4 (2005): 171--200.
link.springer.com. Web.

Billings, Jay J. et al. ``Designing a Component-Based Architecture for
the Modeling and Simulation of Nuclear Fuels and Reactors.'' Proceedings
of the 2009 Workshop on Component-Based High Performance Computing. New
York, NY, USA: ACM, 2009. 6:1--6:4. ACM Digital Library. Web. 25 May
2016. CBHPC '09.

Pizzi, Giovanni et al. ``AiiDA: Automated Interactive Infrastructure and
Database for Computational Science.'' Computational Materials Science
111 (2016): 218--230. ScienceDirect. Web.

The Nek5000 Team. ``Nek5000 \textbar{} A Spectral Element Code for
CFD.'' Nek5000 Website. 22 Oct. 2014. Web. 29 Dec. 2016.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  McAffer, Jeff, Paul VanderLei, and Simon Archer. OSGi and Equinox.
  Boston: Addison-Wesley, 2010. Print.
\item
  McAffer, Jeff, Jean-Michel Lemieux, and Chris Aniszczyk. Eclipse Rich
  Client Platform Second Edition. Boston: Addison-Wesley, 2010. Print.
\end{enumerate}

Gaston, Derek et al. ``MOOSE: A Parallel Computational Framework for
Coupled Systems of Nonlinear Equations.'' Nuclear Engineering and Design
239.10 (2009): n. pag. Print.

Fielding, Roy Thomas. ``Architectural Styles and the Design of
Network-Based Software Architectures.'' University of California,
Irvine, 2000. Print.

Burke, Bill. RESTful Java with JAX-RS. California: O-Reilly, 2010.
Print.

Pontesegger, Christian. ``Eclipse Advanced Scripting Environment.''
N.p., n.d. Web. 1 Jan. 2017.

Billings, Jay Jay. ``Eclipse ICE.'' Project Website. Eclipse ICE. N.p.,
21 Oct. 2016. Web. 1 Jan. 2017.

Tibbitts, Beth R., Gregory R. Watson, and Craig E. Rasmussen. ``An
Integrated Approach to Improving the Parallel Application Development
Process.'' 2009 IEEE International Symposium on Parallel \& Distributed
Processing (IPDPS). Los Alamitos, CA, USA: IEEE Computer Society, 2009.
1--8. IEEE Computer Society. Web.

Billings, Jay Jay. ``Eclipse Advanced Visualization Project.'' Text.
projects.eclipse.org. N.p., 16 Dec. 2015. Web. 1 Jan. 2017.

Billings, Jay Jay et al. ``A Domain-Specific Analysis System for
Examining Nuclear Reactor Simulation Data for Light-Water and
Sodium-Cooled Fast Reactors.'' Annals of Nuclear Energy 85 (2015):
856--868. ScienceDirect. Web.

Billings, Jay Jay. ``The Brand New Neutron Reflectivity Simulator in
Eclipse ICE and What It Took to Make It.'' EclipseCon Europe 2015.
Ludwigsburg Germany: Eclipse Foundation, 2015. Web. 25 May 2016.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Brooks, Christopher. ``Triquetrum: Models of Computation for
  Workflows.'' EclipseCon NA 2016. N.p., 4 Dec. 2015. Web. 1 Jan. 2017.
\item
  Brooks, Christopher, and Jay Jay Billings. ``Introducing Triquetrum, A
  Possible Future for Kepler and Ptolemy II.'' Procedia Computer Science
  80 (2016): 2449--2454. ScienceDirect. Web. International Conference on
  Computational Science 2016, ICCS 2016, 6-8 June 2016, San Diego,
  California, USA.
\end{enumerate}

Humble, T. S. et al. ``An Integrated Programming and Development
Environment for Adiabatic Quantum Optimization.'' Computational Science
\& Discovery 7.1 (2014): 015006. Institute of Physics. Web.

McCaskey. ``Scientific Simulation with Eclipse - From Zero Code to
Running on Lots of Cores in 10 Minutes.'' EclipseCon NA 2016. Reston VA:
Eclipse Foundation, 2015. Web. 26 May 2016.

Pannala, S. et al. ``Multiscale Modeling and Characterization for
Performance and Safety of Lithium-Ion Batteries.'' Journal of Applied
Physics 118.7 (2015): 072017. scitation.aip.org. Web.

``ORNL-QCI/Xacc.'' GitHub. N.p., n.d. Web. 1 Jan. 2017.

Mandal, Nandita et al. ``Integrating Existing Scientific Workflow
Systems: The Kepler/Pegasus Example.'' Proceedings of the 2Nd Workshop
on Workflows in Support of Large-Scale Science. New York, NY, USA: ACM,
2007. 21--28. ACM Digital Library. Web. 20 Dec. 2016. WORKS '07.

Graham, Jonah. ``Eclipse January.'' Text. projects.eclipse.org. N.p., 6
Apr. 2016. Web. 1 Jan. 2017.

\section{Footnotes}\label{footnotes}

1: The authors have identified many different combinations of these
activities that define workflow ``classes.'' When possible, every effort
is made to give the classes funny names such as ``The Re-Run'' or ``The
Graduate Student.''

2: An upcoming update to the API will see the formal introduction of
IWorkflow, IWorkflowTask, and IWorkflowEngine interfaces to bring ICE's
API language closer to other systems. This must be done with care to
satisfy backwards compatibility requirements on the present API.