### 2017-05-23 src/main.tex
Fixed messed up superscripts in affiliations. Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

%%This is a very basic article template.
%%There is just one section and two subsections.
\documentclass{elsart}
%% For Arxiv.org
\pdfoutput=1

\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{tabularx}
\usepackage{supertabular}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage[hyphens]{url}

\begin{document}

\begin{frontmatter}

\title{The Eclipse Integrated Computational Environment}

\author{Jay Jay Billings$^{1,2}$}
\ead{billingsjj@ornl.gov, Twitter: @jayjaybillings}
\author{Andrew R. Bennett$^{1,3}$}
\author{Jordan Deyton$^{1,4}$}
\author{Kasper Gammeltoft$^{1,5}$}
\author{Jonah Graham$^{6}$}
\author{Dasha Gorin$^{1,7}$}
\author{Hari Krishnan$^{8}$}
\author{Menghan Li$^{9}$}
\author{Alexander J. McCaskey$^{1}$}
\author{Taylor Patterson$^{1,10}$}
\author{Robert Smith$^{1}$}
\author{Gregory R. Watson$^{1}$}
\author{Anna Wojtowicz$^{1,11}$}

\address{$^{1}$Computer Science and Mathematics Division, Oak Ridge National
Laboratory, Oak Ridge, TN 37830, USA}
\address{$^{2}$The Bredesen Center for Interdisciplinary Research and Graduate
Education, University of Tennessee, 444 Greve Hall, 821 Volunteer
  Blvd. Knoxville, TN 37996-3394}
\address{$^{3}$University of Washington, Seattle, WA 98105}
\address{$^{4}$General Electric Company, 3200 North Grandview Blvd Waukesha, WI
53188-1678}
\address{$^{5}$Georgia Institute of Technology North Avenue, Atlanta, GA 30332}
\address{$^{6}$Kichwa Coders Ltd. 1 Plomer Green Avenue Downley, High Wycombe
HP13 5LN United Kingdom}
\address{$^{7}$Northwestern University 633 Clark Street Evanston, IL 60208}
\address{$^{8}$Lawrence Berkeley National Laboratory, 1 Cyclotron Rd, Berkeley,
CA 94720}
\address{$^{9}$Department of Computer Science, and Department of Biological
Sciences, Purdue University, West Lafayette, IN 47906} 
\address{$^{10}$Acato Information Management, LLC 114 Union Valley Rd. Oak Ridge,
TN 37830}
\address{$^{11}$Colorado State University Fort Collins, CO 80523}

\begin{abstract}

Problems in modeling and simulation require significantly different
workflow management technologies than standard grid-based workflow
management systems. Computational scientists typically interact with
simulation software in a feedback driven way where solutions and
workflows are developed iteratively and simultaneously. This work
describes common activities in workflows and how combinations of these
activities form unique workflows. It presents the Eclipse Integrated
Computational Environment as a workflow management system and
development environment for the modeling and simulation community.
Examples of the Environment's applicability to problems in energy
science, general multiphysics simulations, quantum computing and other
areas are presented as well as its impact on the community.

\end{abstract}

\begin{keyword}
\sep
\sep 
supercomputing

\vspace{1ex}

\PACS
07.05.Tp % Computer modeling and simulation

\end{keyword}

\end{frontmatter}

\section{Notice of Copyright}\label{notice-of-copyright}

This manuscript has been authored by UT-Battelle, LLC under Contract No.
DEAC05-00OR22725 with the U.S. Department of Energy. The United States
Government retains and the publisher, by accepting the article for
publication, acknowledges that the United States Government retains a
nonexclusive, paid-up, irrevocable, world-wide license to publish or
reproduce the published form of this manuscript, or allow others to do
so, for United States Government purposes. The Department of Energy will
provide public access to these results of federally sponsored research
in accordance with the DOE Public Access Plan
(http://energy.gov/downloads/doe-public-access-plan ).

\input{content}

\bibliographystyle{unsrt}
\bibliography{bib}

\end{document}