### 2019-07-28 src/conclusions.tex
Moved conclusions to separate file.
Signed-off-by: Jay Jay Billings <billingsjj@ornl.gov>

Modeling and simulation workflows for computational scientists differ greatly
from those of experimentalists or those who primarily interact with grid-based workflow
management systems. The Eclipse ICE effort described here has been used to
address interdisciplinary problems in modeling and simulation for energy
science. Of particular interest are the differences in architecture between a
workflow management system focused on modeling and simulation compared with
systems focused on grids. ICE's broad applicability across many topics in energy
science suggests there are opportunities for these systems in general. Finally,
one interesting avenue of future exploration is coupling or integrating ICE with
other workflow tools such as Aiida, Triquetrum, Kepler, and Pegasus, which would
make it possible to combine the best of grid and modeling workflows and
simulation workflows to address greater challenges.